---
title: "GLMs for Car Insurance Pricing: Initial Modeling"
author: "Mick Cooney"
date: "13 July 2016"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: yes
    theme: cerulean
    css: styles.css
  pdf_document: default
---

<!--
(Title:) GLMs for Car Insurance Pricing

Author: Mick Cooney

Date: 2016

Abstract: This document is part of a series investigating the use of
generalised linear models to price car insurance. This document
focuses on the production of some simple GLM models for both a Poisson
and a Gamma process to assess both the frequency and the size of
claims. We will discuss different ways to model claims and then
combine these to produce a price for the premium.


Keywords: car-insurance

-->

```{r knit_opts, include = FALSE}
rm(list = ls())

knitr::opts_chunk$set(tidy       = FALSE
                     ,cache      = FALSE
                     ,fig.height =  8
                     ,fig.width  = 11
                     )

library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(scales)
library(feather)

options(width            = 90)
options(stringsAsFactors = FALSE)

options(datatable.print.nrows      = 10L)
options(datatable.prettyprint.char = 80L)

set.seed(142)

source("custom_functions.R")
```

# Load Data

In the previous document we saved our work as `feather` files and so
we restore those tables now. We will mainly work with the
`policyclaim_dt` table but keep the other two tables in the workspace
as they may be useful later.

```{r load_data, echo=TRUE}
policy1_dt     <- read_feather("data/policy_data.feather")
claim1_dt      <- read_feather("data/claim_data.feather")
policyclaim_dt <- read_feather("data/policyclaim_data.feather")

setDT(policy1_dt)
setDT(claim1_dt)
setDT(policyclaim_dt)

glimpse(policyclaim_dt)
```

We convert the tables to `data.table` as this makes some of the later
work a little faster as we can add columns to tables in-place
efficiently.

For the moment, we are not really building any predictive models, so
we are not doing the standard train/test split. We will approach this
later when building models and using them to produce predictions.

# Modelling Claims

To model the claims in the book of policy we split the claim modelling
into two parts: we first try to model the frequency of claims for each
policy, and then we try to assess the size of the claims. By
multiplying these two together we get a sense of the cost of each
policy.

Due to the highly random nature of claim costs, we expect it to be
much harder to predict the claim size from the covariates. Most of the
predictive power is likely to be in modelling the claims frequency.

We will need to do some further data exploration as part of this work,
but rather than trying to understand the data, the purpose of this is
model-building: we want to get a sense for how the outputs of interest
vary across the different parameters and see if we can improve the
models by using this.

## Modelling the Claim Frequency

The standard method for modelling the frequency of an event is the
*Poisson* process - we assume events occur randomly but distributed as
a Poisson process, and then our predictions affect the parameter of
this process: the *rate* parameter $\lambda$.

The $\lambda$ parameter in a Poisson process is the frequency of the
occurrence for a given unit of time. The most natural unit of time for
us is a year - the standard length of exposure for a policy - and so
if $\lambda = 2$ that means we expect an event to occur twice a year
on average.

Before we model anything with the data, we first take a look at the
overall rate of claims on the book as a whole. This allows us to
anchor our expectations for parameters, not to mention that it serves
as a simple 'sanity' check for the data.

```{r calc_claim_rate, echo=TRUE}
policyclaim_dt %>%
    summarise(claim_rate = sum(claim_count) / sum(exposure))
```

The overall claim rate for the book as a whole is just under 0.07
claims per policy per year.

Breaking this down a little further, we can do a similar calculation
by different combinations of categorical variables, such as by region
or by age category.

```{r claimrate_region, echo=TRUE}
claimrate_region_dt <- policyclaim_dt %>%
    group_by(region) %>%
    summarise(claimrate = sum(claim_count) / sum(exposure))

ggplot(claimrate_region_dt) +
    geom_point(aes(x = region, y = claimrate)) +
    expand_limits(y = c(0, 0.1)) +
    xlab("Region") +
    ylab("Claim Rate") +
    ggtitle("Rate of Claims across Regions")
```

For the purposes of looking at the effect of age on claim rates, we
will look at the categorised data in preference to using it as a
continuous variable. We may use either in the modelling, but for
visualisation of this effect, we use the categories.

```{r claimrate_cat_driverage, echo=TRUE}
claimrate_driverage_dt <- policyclaim_dt %>%
    group_by(cat_driver_age) %>%
    summarise(claimrate = sum(claim_count) / sum(exposure))

ggplot(claimrate_driverage_dt) +
    geom_point(aes(x = cat_driver_age, y = claimrate)) +
    expand_limits(y = c(0, 0.2)) +
    xlab("Driver Age") +
    ylab("Claim Rate") +
    ggtitle("Rate of Claims across Driver Age categories")
```

There is a strong pattern here: younger drivers crash more - not an
unexpected result. We will try the same thing but with the actual
value, just to get a sense of the relationship between age and claim
frequency.

```{r claimrate_driverage_plot, echo = TRUE}
claimrate_driverage_dt <- policyclaim_dt %>%
    mutate(group_age = as.character(driver_age)) %>%
    group_by(group_age) %>%
    summarise(claimrate = sum(claim_count) / sum(exposure))

ggplot(claimrate_driverage_dt) +
    geom_point(aes(x = group_age, y = claimrate)) +
    expand_limits(y = 0) +
    xlab("Driver Age") +
    ylab("Claim Rate") +
    ggtitle("Rate of Claims across Driver Age") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

Treating ages a more continuous variable for grouping does show a
nonlinear form, but for the higher age drivers there is so little data
that we will stick with the categories we have for now and see where
that leads us.

### Poisson Regression

We turn our attention to building a first model for this: we will
construct a GLM for the claim_count and use a Poisson link function to
estimate our regression coefficients. This means we are regressing on
the rate parameter of the Poisson function: our estimates can be
interpreted as directly affected the rate of claims expected.

Before we begin, we do need to deal with the `exposure` variable: not
all policies were in force for 1 year and our model needs to account
for this. Strangely, we have policies in the dataset that had exposure
for longer than a year also - we may end up excluding them later.

To understand this, we will go back to the underlying mathematics of
this process.

Suppose we have $Y_i$, the number of claims from policy $i$ in a
year. Assuming the underlying process is Poisson-distributed, this
means that in units of a year we have

$$ Y_i \sim \mathcal{P}(N_i) $$

where $N_i$ is the underlying rate of claims for policy $i$.

If policy $i$ is only in force for a fraction of the year, $E_i$, then
$N_i$ is unobserveable: we can only observe $Y_i$ during the period of
time $E_i$.

Thus, we have

$$ Y_i \sim \mathcal{P}(\lambda_i \, E_i). $$

For Poisson regression we use a logarithm link function so that

$$ \lambda_i = \exp{(X_i \beta)} $$

and

$$ Y_i \sim \mathcal{P}(e^{X_i \beta + \log E_i}) $$.

For regression purposes, $\log E_i$ is an offset to RHS of the
equation, and we use `offset()` in the formula to account for this.

### A Simple Claim-rate Model

Before we start modeling the rate of claims "in anger", we first get a
sense for how this GLM works - modelling incident rates via a Poisson
process. To that end, we will start with a very simple model and
estimate the claim rate using only the type of fuel the car uses:

```{r claimrate_fuelmodel, echo=TRUE}
model_fuel <- glm(claim_count ~ 0 + fuel + offset(log(exposure))
             ,family = poisson
             ,data = policyclaim_dt)

summary(model_fuel)
```

We now try to understand the meaning of this output.

We have added the $0$ to the formula to prevent the 'Diesel' level for
`fuel` to be added to the intercept - we want two seperate values shown
in the output.

The outputs are on the log scale, if we want to see their 'natural'
values, i.e. the impact they have on the incidence rate, we need to
exponentiate them.

```{r claimrate_fuelmodel_coefs, echo=TRUE}
exp(coef(model_fuel))
```

According to this model, Diesel cars have claims at a rate of 0.0747
per year, whereas Regular fuel cars incur claims at a rate of 0.0652.

As a quick sanity check, we create some sample data and predict the
output. If the data is for a year's exposure with both a Diesel and
Regular car, we expect to get similar answers as above:

```{r claimrate_fuelmodel_predict, echo=TRUE}
test_dt <- data.table(fuel     = c("Diesel", "Regular")
                     ,exposure = 1)

predict(model_fuel, newdata = test_dt, type = 'response')
```

We see that the predicted claim rates are the same as the coefficients
of the model, matching what we would expect.

How does this compare to the observed claim rate observed by splitting
across fuel type?

```{r claimrate_fuel_observed, echo=TRUE}
policyclaim_dt %>%
    group_by(fuel) %>%
    summarise(claim_rate = sum(claim_count) / sum(exposure))
```

We recover the same claim rates by fuel using all the above
approaches. This makes sense, as we are doing the same thing in
different ways, so it is good that we get the same answer each time!


### Building the GLM

We start with a simple GLM with a small number of covariates (using
some of our new engineered features):

```{r claimrate_model_1, echo=TRUE}
model1 <- glm(claim_count ~ cat_driver_age + density + agg_region +
                            offset(log(exposure))
             ,family = poisson
             ,data = policyclaim_dt)

summary(model1)
```

Not that since we have a number of categorical variables, the first
level of each variable is subsumed by the intercept.

This explains why all the `cat_driver_age` levels in the summary are
negative. Younger drivers have higher claim rates and so being in a
different age category tends to reduce the rate of claims made.

The above model has a single continuous variable: `density` and it is
quite possible that assuming a linear relationship between this and
the claim rate is flawed, so we will remove it and refit the model to
see the effect.

```{r claimrate_model_2, echo=TRUE}
model2 <- glm(claim_count ~ cat_driver_age + agg_region +
                            offset(log(exposure))
             ,family = poisson
             ,data = policyclaim_dt)

summary(model2)
```

This model does not fit as well - there was information in `density`
that is not captured by `region`.

#### Adding Vehicle Details

We also need to take into account the type of vehicle being
insured. It seems sensible that different car types will have
different effects on the likelihood of a claim. The dataset has a
number of variables related to the vehicle:

* `car_age`
* `power`
* `brand`
* `fuel`

We will try adding all the variables (using our modified versions
where appropriate):

```{r claimrate_model_3, echo=TRUE}
model3 <- glm(claim_count ~ cat_driver_age + agg_region +
                            cat_car_age + agg_power_2 + brand + fuel +
                            offset(log(exposure))
             ,family = poisson
             ,data = policyclaim_dt)

summary(model3)
```

The AIC for this model is smaller, matching our intuition that vehicle
data has a strong effect on the rate of claims for a policy.

Our current aim is merely to get a feel for the models, so we will not
pay too much attention to model assessment or validation. For
starters, we have done little pruning of the data and have no created
any sort of holdout set, and such a process could easily change our
inferences!

Instead, we leave it there and instead turn our attention to modelling
the size of the claims.


## Modelling Claim Size

We turn our attention to modelling the size of the claim, given that a
claim has occurred. We use the individual claim data for this, but
will use variables from the claim data also.




# Model Assessment