---
title: "GLMs for Car Insurance Pricing: Initial Modeling"
author: "Mick Cooney"
date: "13 July 2016"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: yes
    theme: cerulean
    css: styles.css
  pdf_document: default
---

<!--
(Title:) GLMs for Car Insurance Pricing

Author: Mick Cooney

Date: 2016

Abstract: This document is part of a series investigating the use of
generalised linear models to price car insurance. This document
focuses on the production of some simple GLM models for both a Poisson
and a Gamma process to assess both the frequency and the size of
claims. We will discuss different ways to model claims and then
combine these to produce a price for the premium.


Keywords: car-insurance

-->

```{r knit_opts, include = FALSE}
rm(list = ls())

knitr::opts_chunk$set(tidy       = FALSE
                     ,cache      = FALSE
                     ,fig.height =  8
                     ,fig.width  = 11
                     )

library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(scales)
library(feather)

options(width            = 90)
options(stringsAsFactors = FALSE)

options(datatable.print.nrows      = 10L)
options(datatable.prettyprint.char = 80L)

set.seed(142)

source("custom_functions.R")
```

# Load Data

In the previous document we saved our work as `feather` files and so
we restore those tables now. We will mainly work with the
`policyclaim_dt` table but keep the other two tables in the workspace
as they may be useful later.

```{r load_data, echo=TRUE}
policy1_dt     <- read_feather("data/policy_data.feather")
claim1_dt      <- read_feather("data/claim_data.feather")
policyclaim_dt <- read_feather("data/policyclaim_data.feather")

setDT(policy1_dt)
setDT(claim1_dt)
setDT(policyclaim_dt)

glimpse(policyclaim_dt)
```

We convert the tables to `data.table` as this makes some of the later
work a little faster as we can add columns to tables in-place
efficiently.
