
---
title: "GLMs for Car Insurance Pricing: Model Assessment"
author: "Mick Cooney"
date: "13 July 2016"
output:
  html_document:
    toc: true
    number_sections: true
    fig_caption: yes
    theme: cerulean
    css: styles.css
  pdf_document: default
---

<!--
(Title:) GLMs for Car Insurance Pricing

Author: Mick Cooney

Date: 2016

Abstract: This document is part of a series investigating the use of
generalised linear models to price car insurance. This document
assesses the pricing models built previously and uses MonteCarlo
simulation to discover the range of possible outcomes that are
possible assuming we get the data correct. The main focus is to get
assess the distribution of possible outcomes.

Keywords: car-insurance pricing assessment

-->

```{r knit_opts, include = FALSE}
rm(list = ls())

knitr::opts_chunk$set(tidy       = FALSE
                     ,cache      = FALSE
                     ,fig.height =  8
                     ,fig.width  = 11
                     )

library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(scales)
library(feather)
library(poweRlaw)


options(width            = 90)
options(stringsAsFactors = FALSE)

options(datatable.print.nrows      = 10L)
options(datatable.prettyprint.char = 80L)

source("custom_functions.R")
```

# Set Parameters and Load Data

Before we begin we configure some parameters needed for the
simulations.

```{r config_params, echo=TRUE}
set.seed(242)

n_sim <- 100

largeloss_threshold <- 5000
```

We load the various datasets, the model fits and the price calculation.

```{r load_data, echo=TRUE}
trainpolicy_dt <- read_feather("data/train_data.feather")
validpolicy_dt <- read_feather("data/valid_data.feather")
testpolicy_dt  <- read_feather("data/test_data.feather")


setDT(trainpolicy_dt)
setDT(validpolicy_dt)
setDT(testpolicy_dt)

glimpse(trainpolicy_dt)
glimpse(validpolicy_dt)
glimpse(testpolicy_dt)

claimrate_model_glm <- readRDS("data/claim_model.rds")
claimsize_model_glm <- readRDS("data/size_model.rds")

premium_quoter <- readRDS("data/premium_quoter.rds")
```

In this document we focus assessing our pricing model. No model is
perfect, and it is important to understand how a pricing model may
fail and what risks are posed.

To start with, we make a very strong and generous assumption: we
assume our pricing model is correct and the risks assumed by the
business are represented correctly by the models.

This assumption is absurd, we already know the model has flaws, and
spent part of the last document working on them. That said, it makes
for a easy and frictionless starting point, and should help illustrate
the extent of tail risk we have in the model.

# Evaluating the Model with Same Parameters

The idea is to use the same parameters used to price the policies to
simulate claim data. For each simulation iteration and policy, we
perform a number of steps:

* Use the Poisson model to estimate claim counts for the policy
* Use the Gamma model to estimate the size of the claims
* Use the largeloss claim frequency to estimate the count of large claims
* Use the powerlaw calculation to simulate the size of the large claims

Note that for performance reasons the logical steps outlined above may
be implemented differently without changing the outcome of the
simulation overall.

Having done the above for every policy and claim, we run multiple
MonteCarlo simulations to estimate the distribution of possible
outcomes *conditional on the model being correct*.

This approach may seem pointless - using a model to validate itself -
but it is not. It gives us a sense of the best case scenario from a
probabilistic perspective.


```{r model_assess_simple, echo=TRUE}

create_mc_simulator <- function(claimrate_model_glm
                               ,claimsize_model_glm
                               ,largeloss_sf
                               ,policydata_dt) {

    claim_rate <- predict(claimrate_model_glm, newdata = policydata_dt, type = 'response')

    run_mc_sim <- function() {
        return(1)
    }

    return(run_mc_sim)
}


```